SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-12-22 maintex.tex
updated figures

\begin{abstract}
We compare the radial profiles of the specific star formation rate (sSFR) in a sample of 175 star-forming galaxies in close pairs with those of mass-matched \textbf{control} galaxies in the SDSS-IV MaNGA survey. We find that the sSFR is centrally enhanced (within one effective radius) in interacting galaxies by $\sim$0.25 dex and that there is a weak sSFR suppression in the outskirts of the galaxies of $\sim$0.1 dex. We stack the differences profiles for galaxies in five stellar mass bins between \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass. The same result is obtained when the comparison galaxies are matched to each paired galaxy in both stellar mass and redshift. In addition, we find that that the sSFR enhancement is elevated in pairs with nearly equal masses and closer projected separations, in agreement with previous work based on single-fiber spectroscopy. We also find that the sSFR offsets in the outskirts of the paired galaxies are dependent on whether the galaxy is the more massive or less massive companion in the pair. The more massive companion experiences zero to a positive sSFR enhancement while the less massive companion experiences sSFR suppression in their outskirts.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{color-mag.pdf}
\caption[]{Color-magnitude diagram for MaNGA galaxies. The grey circles represent the whole MaNGA sample and the color scale reflects the local density around each data point in this color-magnitude plane, as indicated by the color bar on the top. The {\it blue} diamonds represent the star forming paired galaxies (by the BPT and \ewha\ cuts), the {\it red} diamonds represent the paired galaxies which are classified as retired (by the \ewha\ cut), and the {\it green} diamonds represent the paired galaxies containing an AGN (as identified by the BPT diagnostic). From top to bottom, the dashed lines divide the sample into red sequence, green valley, and blue cloud. The star-forming galaxy sample used in this paper are the galaxies below the lower dividing line.}
\label{fig:cmd}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

From the NSA catalog's 641,409 galaxies, we find 492 galaxies which are paired to MaNGA targets. After restricting the sample to SFGs, we have another 133 MaNGA targets with paired galaxies outside of the IFU. MaNGA targets which have both an inside-IFU and an outside-IFU pair are left to the inside-IFU sample. This give us a total of 175 close pairs.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=\linewidth]{ssfr_comb.pdf}
\caption[]{The log(sSFR) as a function of galactocentric radius for control galaxies (\textbf{Left}) and galaxy pairs (\textbf{Middle}). The difference between the profiles of the paired galaxies and the control galaxies are shown in the \textbf{Right} panel. The dashed black profile represents the mean of the difference profiles. The colors of the profiles represent the mass range of the selected galaxies which is given in the legend along with the number of galaxies in that mass bin in parantheses. The highlighted region around the profiles represent the standard error of the mean of the data at the given radius interval. }
\label{fig:ssfr_prof}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{ssfr_mass.pdf}
\caption[]{The \textbf{Left} panel shows the $\Delta$log(sSFR) profiles where the difference profiles are constructed from the difference between the paired galaxy profiles and a set of 20 control galaxies. The profiles are split into five different stellar mass bins and the highlighted region about the profiles represent the standard error of mean of the profile. The black dashed line represents the mean profile between the four difference mass ranges. The number of paired galaxies in each mass range is given in the legend in parentheses. In the \textbf{Right} panel the nuclear $\Delta$log(sSFR) are shown. The black squares are the mean values within a stellar mass bin (where the size of the bins are shown the the horizontal error bars). The vertical error bars on the black squares represent the standard deviation within the bin. The horizontal, dashed black line represents the median central enhancement of the pair sample. Galaxies in the outside-IFU (Blue) and inside-IFU (Red) samples are separately depicted.}
\label{fig:ssfr_mass}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{sep_hist.pdf}
\caption[]{The projected separation distribution of the whole sample (blue) against the projected separation distribution for the highest mass galaxies, \logm\ $=$ 11.0$-$11.5 (Red). }
\label{fig:sep_hist}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Tailored-Controls Difference Profiles}\label{sec:tailored}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{ssfr_dm.pdf}
\caption[]{Same as Figure \ref{fig:ssfr_mass} except the difference profiles are split by the mass ratio of the pair. The inside-IFU sample is left out of this analysis because we do not yet have reliable measurements of their total stellar mass ratios.}
\label{fig:ssfr_dm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the second method, we match each paired galaxy to a set of 20 control galaxies of similar stellar masses and redshifts, as described in \S~\ref{sec:control}. We then take the median of the azimuthally averaged profiles of the tailored control sample. Finally, we obtain the $\Delta\log$sSFR profile for each of the 175 paired galaxies by calculating the difference between its sSFR profile and the median profile of its control sample.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{ssfr_sep.pdf}
\caption[]{Same as Figure \ref{fig:ssfr_mass} except the difference profiles are split by projected separation. }
\label{fig:ssfr_sep}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=3in]{prof_comp.pdf}
\caption[]{The mean radial profile of $\Delta$log(sSFR) between pairs and controls of this work with the tailored control method ({\it Black}) compared against those of \citet{Pan:2019} ({\it Red}), a pair sample, and \citet{Thorp:2019} ({\it Blue}), a post-merger sample.}
\label{fig:prof_comp}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=3in]{nuc_sep.pdf}
\caption[]{(\textbf{Top}) $\Delta$log(SFR) over projected separation from this study ({\it Black}), \citet{Ellison:2008} ({\it Red}), \citet{Scudder:2012} ({\it Blue}), \citet{Patton:2013} ({\it Green}), and \citet{Bustamante:2020} ({\it Purple}). (\textbf{Bottom}) $\Delta$log(sSFR) over projected separation from this study ({\it Black}), \citet{Li:2008} ({\it Red}), and \citet{Patton:2020} ({\it Blue}).}
\label{fig:nuc_sep}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\begin{abstract}
We compare the radial profiles of the specific star formation rate (sSFR) in a sample of \textbf{169} star-forming galaxies in close pairs with those of mass-matched \textbf{control} galaxies in the SDSS-IV MaNGA survey. We find that the sSFR is centrally enhanced (within one effective radius) in interacting galaxies by $\sim$0.25 dex and that there is a weak sSFR suppression in the outskirts of the galaxies of $\sim$0.1 dex. We stack the differences profiles for galaxies in five stellar mass bins between \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass. The same result is obtained when the comparison galaxies are matched to each paired galaxy in both stellar mass and redshift. In addition, we find that that the sSFR enhancement is elevated in pairs with nearly equal masses and closer projected separations, in agreement with previous work based on single-fiber spectroscopy. {\it We also find that the sSFR offsets in the outskirts of the paired galaxies are dependent on whether the galaxy is the more massive or less massive companion in the pair. The more massive companion experiences zero to a positive sSFR enhancement while the less massive companion experiences sSFR suppression in their outskirts.} \textbf{Our results indicate that merger induced gas-inflows are responsible for elevating the central sSFR in galaxy pairs. }
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{fig/bpt-cmd.pdf}
\caption[]{Color-magnitude diagram for MaNGA galaxies. The grey circles represent the whole MaNGA sample and the color scale reflects the local density around each data point in this color-magnitude plane, as indicated by the color bar on the top. The {\it blue} diamonds represent the star forming paired galaxies (by the BPT and \ewha\ cuts), the {\it red} diamonds represent the paired galaxies which are classified as retired (by the \ewha\ cut), and the {\it green} diamonds represent the paired galaxies containing an AGN (as identified by the BPT diagnostic). From top to bottom, the dashed lines divide the sample into red sequence, green valley, and blue cloud. The star-forming galaxy sample used in this paper are the galaxies below the lower dividing line.}
\label{fig:cmd}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

From the NSA catalog's 641,409 galaxies, we find 492 galaxies which are paired to MaNGA targets. After restricting the sample to SFGs, we have another 133 MaNGA targets with paired galaxies outside of the IFU. MaNGA targets which have both an inside-IFU and an outside-IFU pair are left to the inside-IFU sample. This give us a total of \textbf{169} close pairs.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=\linewidth]{fig/ssfr_comb.pdf}
\caption[]{The log(sSFR) as a function of galactocentric radius for control galaxies (\textbf{Left}) and galaxy pairs (\textbf{Middle}). The difference between the profiles of the paired galaxies and the control galaxies are shown in the \textbf{Right} panel. The dashed black profile represents the mean of the difference profiles. The colors of the profiles represent the mass range of the selected galaxies which is given in the legend along with the number of galaxies in that mass bin in parantheses. The highlighted region around the profiles represent the standard error of the mean of the data at the given radius interval. }
\label{fig:ssfr_prof}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{fig/ssfr_mass.pdf}
\caption[]{The \textbf{Left} panel shows the $\Delta$log(sSFR) profiles where the difference profiles are constructed from the difference between the paired galaxy profiles and a set of 20 control galaxies. The profiles are split into five different stellar mass bins and the highlighted region about the profiles represent the standard error of mean of the profile. The black dashed line represents the mean profile between the four difference mass ranges. The number of paired galaxies in each mass range is given in the legend in parentheses. In the \textbf{Right} panel the nuclear $\Delta$log(sSFR) are shown. The black squares are the mean values within a stellar mass bin (where the size of the bins are shown the the horizontal error bars). The vertical error bars on the black squares represent the standard deviation within the bin. The horizontal, dashed black line represents the median central enhancement of the pair sample. Galaxies in the outside-IFU (Blue) and inside-IFU (Red) samples are separately depicted.}
\label{fig:ssfr_mass}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{fig/sep_hist.pdf}
\caption[]{The projected separation distribution of the whole sample (blue) against the projected separation distribution for the highest mass galaxies, \logm\ $=$ 11.0$-$11.5 (Red). }
\label{fig:sep_hist}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Tailored-Controls Difference Profiles}\label{sec:tailored}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{fig/ssfr_dm.pdf}
\caption[]{Same as Figure \ref{fig:ssfr_mass} except the difference profiles are split by the mass ratio of the pair. The inside-IFU sample is left out of this analysis because we do not yet have reliable measurements of their total stellar mass ratios.}
\label{fig:ssfr_dm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the second method, we match each paired galaxy to a set of 20 control galaxies of similar stellar masses and redshifts, as described in \S~\ref{sec:control}. We then take the median of the azimuthally averaged profiles of the tailored control sample. Finally, we obtain the $\Delta\log$sSFR profile for each of the \textbf{169} paired galaxies by calculating the difference between its sSFR profile and the median profile of its control sample.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{fig/ssfr_sep.pdf}
\caption[]{Same as Figure \ref{fig:ssfr_mass} except the difference profiles are split by projected separation. }
\label{fig:ssfr_sep}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=3in]{fig/prof_comp.pdf}
\caption[]{The mean radial profile of $\Delta$log(sSFR) between pairs and controls of this work with the tailored control method ({\it Black}) compared against those of \citet{Pan:2019} ({\it Red}), a pair sample, and \citet{Thorp:2019} ({\it Blue}), a post-merger sample.}
\label{fig:prof_comp}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=3in]{fig/nuc_sep.pdf}
\caption[]{(\textbf{Top}) $\Delta$log(SFR) over projected separation from this study ({\it Black}), \citet{Ellison:2008} ({\it Red}), \citet{Scudder:2012} ({\it Blue}), \citet{Patton:2013} ({\it Green}), and \citet{Bustamante:2020} ({\it Purple}). (\textbf{Bottom}) $\Delta$log(sSFR) over projected separation from this study ({\it Black}), \citet{Li:2008} ({\it Red}), and \citet{Patton:2020} ({\it Blue}).}
\label{fig:nuc_sep}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
